---
output: github_document
bibliography: vignettes/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# MGPMMammals

This R-package contains the data and the R-code for the analysis of the mammal dataset in the research article [@Mitov:2019a] (see References).

## Installation

The easiest way to install this package is to use the function `install_github()` from the R-package devtools. The following command will install the most up-to-date version of the master branch from the github repository of the package. 

```{r, eval=FALSE}
devtools::install_github("venelin/MGPMMammals")
```


## License 

All R and Rmd code files in MGPMMammals are distributed under the terms of the 
[Creative Commons Attribution 4.0 International License (CC BY 4.0)](https://creativecommons.org/licenses/by/4.0/legalcode). 

Special terms apply to the following data-files and objects:

- MGPMMammals/data/tree.rda and ALL identical or derived tree objects
describing the mammal phylogenetic tree (i.e. objects of S3 class phylo),
that are included in the data-files MGPMMammals/data/fitMappings*,
MGPMMammals/data/bestFitToDataWithSEs.rda,
MGPMMammals/data/bestModelToDataWithoutSEs.rda,
MGPMMammals/fits_bootstrap_MGPM_A_F_best_clade_RR_HD.rda. All of these tree
objects are derived from the Mammal phylogenetic tree published in [@BinindaEmonds:2007bk]
These data-objects are included in MGPMMammals solely for the purpose of reproducibility; 
Hence, the author of MGPMMammals does not hold a copyright over any of these data-objects.
If you reuse any of these tree objects in a publication, please cite this publication as explained in the output of the following R-coce `print(citation("MGPMMammals"), bibtex = TRUE)` (see next section).
- MGPMMammals/data-raw/brain_body_database_v2.txt, 
MGPMMammals/data-raw/jeb_2491_sm_tables1.xlsx, 
MGPMMammals/data/data.BoddyEtAl.2.rda, MGPMMammals/data/values.rda,
MGPMMammals/data/SEs.rda, MGPMMammals/data/valuesPPCA.rda,
MGPMMammals/data/SEsPPCA.rda as well as ALL identical or derived data-objects
describing the mammal taxonomy and brain and body-mass trait, which are
included in in the data-files located in MGPMMammals/data/fitMappings*,
MGPMMammals/data/bestFitToDataWithSEs.rda,
MGPMMammals/data/bestModelToDataWithoutSEs.rda,
MGPMMammals/fits_bootstrap_MGPM_A_F_best_clade_RR_HD.rda.
All of these data objects are derived from the data published in [@Boddy:2012dd] and in the
accompanying data-package [@Boddy:2012bd]. If you reuse any of these data objects in a publication, please cite the above article as explained in the following section.

## Citing MGPMMammals
Please, read the following:
```{r}
print(citation("MGPMMammals"), bibtex = TRUE)
```

## Further information
For additional information about the data objects and scripts in the package, read [SI Appendix, Section H](https://www.pnas.org/lookup/suppl/doi:10.1073/pnas.1813823116/-/DCSupplemental) from [@Mitov:2019a].

# References
