---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# MGPMMammals

Contains the code and fitted models of the mammal data analysis for 
  the research article entitled "Automatic Generation of Evolutionary Hypotheses 
  using Mixed Gaussian Phylogenetic Models".

## Installation

This package contains large data files and is not intended for release on CRAN. The recommended way to use this package is to follow these steps:

* Clone the package repository from github using the command (type the following three commands in your terminal / shell):
```
git clone https://github.com/venelin/PCMFit.git
```

* Enable git large file system (LFS) in your working copy:
```
git lfs install
```

* Pull the binary data from the LFS
```
git pull
```

* Install the package from your local working copy (type the following command in your R interpreter). Note that the directory MGPMMammals should be in your current directory from which R is running:
```{r, eval = FALSE}
install.packages("MGPMMammals", repos=NULL, type="source")
```

